---
title: "Assessment_3"
author: "Emelia Appiah"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 7,  
  fig.height = 4
  )
```

```{r load-package libraries, inlude = FALSE}
library(pacman)
pacman::p_load(
  rio, here, tidyverse,knitr,kableExtra, skimr,
  formatR, gridExtra, janitor, cowplot, ggpubr, scales
)

```

```{r import-data}
nhanes_df <- import(here("data", "cleaned_NHANES.csv"))

diet_df <- import(here("data", "diet.csv"))

```


# Question 1: Reproducing Figures
The cleaned labels for both gender and ethnicity are used instead of the codes in the raw dataset. 

To keep the same colour assignment as the original figures `gender` was ordered with `Male` first. (code `1` = male, code `2` = female).
The ethnicity categories were also ordered following the original `RIDRETH3` codes (1 = Mexican, 2 = Other Hispanic, 3 = White, 4 = Black, 6 = Asian, 7 = Other).

The age histogram in the original figure only contained adults, so the data was restricted to participants aged 20 and older. Bins were set with `binwidth = 5`and `boundary = 20`, so each bin covers exactly five years of age.
```{r recreating age-histogram and Ethnicity-histogram}
#Age histogram
age_hist <- nhanes_df %>%
  filter(age >= 20) %>%
   mutate(gender = factor(gender, levels = c("Male", "Female"))) %>%
  ggplot(aes(x = age, fill = gender)) +
  geom_histogram(
    binwidth = 5,
    boundary = 20,
    position = "dodge"
  )

#Ethnicity column Barchart
ethnicity_bar <- nhanes_df%>%
    mutate(
    ethnicity_2 = factor(
      ethnicity_2,
      levels = c(
        "Mexican",
        "Other Hispanic",
        "White",
        "Black",
        "Asian",
        "Other"
      )
    ),
    gender = factor(gender, levels = c("Male", "Female"))
  ) %>%
  ggplot(aes(x = ethnicity_2, fill = gender)) +
  geom_bar(position = "fill")

age_hist
ethnicity_bar

```
## Combining the figures
```{r combining figures using cowplot}
combined_figures_cowplot <- plot_grid(age_hist, ethnicity_bar, ncol = 1, 
          labels = c("A", "B"), align = "v")
annotate_figure(
  combined_figures_cowplot,
  top = text_grob(
    "Distribution of Age and Ethnicity Among NHANES Participants (by cowplot)",
    face = "bold",
    size = 14
  )
)
```


```{r combining figures using ggpubr}
combined_figures_ggarange <- ggarrange(age_hist, ethnicity_bar,
          ncol = 1, labels = c("A", "B"),
          common.legend = TRUE, legend = "right")
annotate_figure(
  combined_figures_ggarange,
  top = text_grob(
    "Distribution of Age and Ethnicity Among NHANES Participants (by ggarange)",
    face = "bold",
    size = 14
  )
)
```
**Comparing `plot_grid()` and `ggarrange()`.** 
For these two plots, the `ggarrange()` layout looks cleaner and is the better choice although both functions stack the two plots and add panel labels. 

With the `align = "v"` the `plot_grid()` output lines up the two panels so their plotting areas share the same left edge. However,`plot_grid()` keeps each plot's own legend, so the same `gender` legend is drawn twice, which wastes space and adds nothing. 

`ggarrange()` on the other hand has a built-in `common.legend = TRUE` option that draws one shared legend for the whole figure. Because both plots use the same fill variable with the same colors, a single legend is all the reader needs, and the panels get more width. 

# Question 2: Visualizing key characteristics
```{r summary statistics +missing values}
nhanes_df %>%
  summarise(across(c(age, gender, ethnicity_1, ethnicity_2),
                   ~ sum(is.na(.x)))) %>%
  knitr::kable(caption = "Number of missing values per variable")

```
None of the four variables had missing values, so every one of the
`r comma(nrow(nhanes_df))` participants is included in the plots below.

## Age distribution
```{r age distirbution}
age_plot <- ggplot(nhanes_df, aes(x = age)) +
  geom_histogram(binwidth = 5, boundary = 0, closed = "left",
                colour = "white") +
  labs(title = "Age distribution of NHANES participants",
       x = "Age (years)", y = "Number of participants") +
  theme_minimal()

age_plot

ggsave("age_plot.png",
  plot = age_plot,
  path = here("figures"),
  width = 7,
  height = 4
)
```
Age ranged from `r min(nhanes_df$age)` to `r max(nhanes_df$age)` years, with a median of `r median(nhanes_df$age)` years. The distribution is right-skewed: children and teenagers make up a large share of the sample (`r comma(sum(nhanes_df$age < 18))` participants under 18), and number of participants get smaller in the older age groups. 

## Gender
```{r ender distribution}
gender_plot <- ggplot(nhanes_df, aes(x = gender, fill = gender)) +
  geom_bar() +
  labs(
    title = "GenderDistribution of NHANES participants",
    x = "Gender",
    y = "Number of participants"
  )

gender_plot

ggsave("gender_plot.png",
  plot = gender_plot,
  path = here("figures"),
  width = 7,
  height = 4
)

```
The sample is close to balanced, with `r comma(sum(nhanes_df$gender == "Female"))` females (`r percent(mean(nhanes_df$gender == "Female"), 0.1)`) and `r comma(sum(nhanes_df$gender == "Male"))` males. There were no missing values.

## Ethnicity
```{r Ethnicity distribution}
#ethnicity1
ethnicity_1_plot <- ggplot(nhanes_df, aes(
    x = fct_infreq(ethnicity_1),
    fill = ethnicity_1
  )) +
  geom_bar() +
  labs(
    title = "Distribution of Ethnicity",
    x = "Ethnicity",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ethnicity 2
ethnicity_2_plot <- ggplot(nhanes_df, aes(x = fct_infreq(ethnicity_2), fill = ethnicity_2)) +
  geom_bar() +
  labs(
    title = "Distribution of Ethnicity",
    x = "Ethnicity",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ethnicity_1_plot
ethnicity_2_plot

ggsave("ethnicity_1_plot.png", 
      plot = ethnicity_1_plot,
      path = here("figures"),
      width = 7,
      height = 4 
)

ggsave("ethnicity_2_plot.png", 
      plot = ethnicity_2_plot,
      path = here("figures"),
      width = 7,
      height = 4 
)

```
White is the largest group (`r comma(sum(nhanes_df$ethnicity_2 == "White"))` participants), followed by Black and Mexican participants. The two variables only differ in one place: `ethnicity_1` has five categories, while `ethnicity_2` has six, because it separates **Asian** participants out of the **Other** category.
```{r Compare-ethnicity}
table(ethnicity_1 = nhanes_df$ethnicity_1,
      ethnicity_2 = nhanes_df$ethnicity_2) %>%
  knitr::kable(caption = "Cross-tabulation of the two ethnicity variables")
```
The cross-tabulation shows the two variables agree for every participant outside the "Other" group. In `ethnicity_1`, "Other" holds
`r comma(sum(nhanes_df$ethnicity_1 == "Other"))` people, of whom
`r comma(sum(nhanes_df$ethnicity_2 == "Asian"))`
(`r percent(sum(nhanes_df$ethnicity_2 == "Asian") /  sum(nhanes_df$ethnicity_1 == "Other"), 1)`) are identified as Asian in `ethnicity_2`.

**Decision:**   i  will keep `ethnicity_2` because it contains all the information in `ethnicity_1` plus a separate Asian category, so nothing is lost by dropping `ethnicity_1`. Lumping Asian participants into a mixed "Other" group hides a large, identifiable group and makes "Other" hard to interpret, which matters when the outcome (here, blood pressure) may differ between groups.
```{r drop ethnicity-1}
nhanes_df <- nhanes_df %>%
  select(-ethnicity_1)

```

# Question 3: Improving a ggplot figure
## Critique and Improvement of the Original Figure

The original figure showed the absolute weight of each participant over the 16-week study period. Although it provided information about individual weight trajectories, it was difficult to determine how much each participant's weight changed relative to their starting point. Because participants began the study at different weights, comparing the lines directly did not clearly show the magnitude of weight loss or gain. For example, a participant who started at a higher weight may appear higher on the graph even if their percentage or absolute weight change was similar to that of another participant.

The original figure also used a different color for each participant, resulting in a large legend with 19 different participant labels. This made it difficult to distinguish the lines and required a reader to repeatedly refer to the legend to identify participants. The many overlapping lines further contributed to visual clutter. In addition, the original graph did not provide a clear reference point for identifying whether a participant gained or lost weight relative to their baseline.

The original figure also lacked an informative overall title or subtitle to explain the purpose of the visualization and guide a reader's interpretation. Although the axes are labeled "Weight" and "Week," the reader must infer that the figure is intended to show changes in participants' weight during the diet study. A descriptive title and subtitle would provide context and communicate the main purpose of the figure more clearly.

To improve the visualization, the change in weight for each participant was calculated from the baseline weight at week 0 and the weight at each subsequent week using `Weight - baseline`. This placed all participants on a common baseline of zero, making the direction and magnitude of weight change easier to compare. A dashed horizontal line at zero was added as a visual reference, so values below zero represent weight loss and values above zero represent weight gain.

`facet_wrap()` was used to create a separate panel for each participant. This removed the overlapping trajectories from a single plot and makes each participant's pattern easier to follow. Rather than assigning a separate color to every participant, participants were grouped into three response categories: "Lost weight," "No change," and "Gained weight." This reduced the number of colors and gave the visualization a clearer story about the overall response patterns.

A descriptive title, "Change in Weight From Baseline," and a subtitle explaining that the weight change is measured relative to each participant's week-0 weight. These additions provided immediate context and made the purpose of the visualization clear before a reader examined the individual trajectories.

Finally, `scale_x_continuous()` was used to display the study weeks at 0, 4, 8, 12, and 16 and `scale_y_continuous()` to provide consistent weight-change reference points. Overall, these changes made the figure less cluttered, easier to interpret, and more directly aligned with the goal of showing how each participant's weight changed from their baseline.

## Improved figure

```{r improved plot for diet change}
diet_change <- diet_df %>%
  group_by(Participant) %>%
  mutate(
    baseline = Weight[Week == 0],
    change = Weight - baseline,
   pct_change = (change / baseline) * 100
 ) %>%
  ungroup() %>%
  group_by(Participant) %>%
  mutate(
    response = case_when(
      last(change) < 0 ~ "Lost weight",
      last(change) > 0 ~ "Gained weight",
      TRUE ~ "No change"
    )
  ) %>%
  ungroup()

ggplot(diet_change, aes(x = Week, y = change, color = response)) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "gray40"
  ) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ Participant) +
  
    # Control x-axis
  scale_x_continuous(
    breaks = c(0, 4, 8, 12, 16)
  ) +
  
  # Control y-axis
  scale_y_continuous(
    breaks = c(-20, -15, -10, -5, 0, 5)
  ) +
  labs(
    title = "Change in Weight From Baseline",
    subtitle = "Weight change relative to each participant's week-0 weight",
    x = "Week",
    y = "Change in Weight",
    color = "Response"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```
## Summary of Weight Change by Participants
At week `r max(diet_change$Week)`, participants were grouped according to their weight-change response. The **lost weight** group included `r sum(diet_change$Week == max(diet_change$Week) & diet_change$response == "Lost weight")` participants, with a mean weight change of `r round(mean(diet_change$change[diet_change$Week == max(diet_change$Week) & diet_change$response == "Lost weight"], na.rm = TRUE), 1)` units and a mean percentage change of `r round(mean(diet_change$pct_change[diet_change$Week == max(diet_change$Week) & diet_change$response == "Lost weight"], na.rm = TRUE), 1)`%.

The **gained weight** group included `r sum(diet_change$Week == max(diet_change$Week) & diet_change$response == "Gained weight")` participants, with a mean weight change of `r round(mean(diet_change$change[diet_change$Week == max(diet_change$Week) & diet_change$response == "Gained weight"], na.rm = TRUE), 1)` units and a mean percentage change of `r round(mean(diet_change$pct_change[diet_change$Week == max(diet_change$Week) & diet_change$response == "Gained weight"], na.rm = TRUE), 1)`%.

The **no change** group included `r sum(diet_change$Week == max(diet_change$Week) & diet_change$response == "No change")` participants, with a mean weight change of `r round(mean(diet_change$change[diet_change$Week == max(diet_change$Week) & diet_change$response == "No change"], na.rm = TRUE), 1)` units and a mean percentage change of `r round(mean(diet_change$pct_change[diet_change$Week == max(diet_change$Week) & diet_change$response == "No change"], na.rm = TRUE), 1)`%.

Overall, the summary showed that weight loss was the most common response at week `r max(diet_change$Week)`, with the lost-weight group showing a substantially larger average change from baseline than the gained-weight group.

```{r ex3-summary}
diet_change %>%
  filter(Week == max(Week)) %>%
  group_by(response) %>%
  summarise(
    n = n(),
    mean_weight_change = round(mean(change, na.rm = TRUE), 1),
    mean_change_pct = round(mean(pct_change, na.rm = TRUE), 1)
  ) %>%
  knitr::kable(
    col.names = c(
      "Response group",
      "n",
      "Mean weight change",
      "Mean change (%)"
    ),
    caption = "Weight change at week 16 by response group"
  )
```
# Question 4: Exploring relationships through visualizations

```{r preparing SBP}
nhanes_df <- nhanes_df %>%
  mutate(
    avg_sbp = if_else(
      rowSums(!is.na(cbind(systolic_bp_1, systolic_bp_2, systolic_bp_3, systolic_bp_4))) == 0,
      NA_real_,
      rowMeans(
        cbind(systolic_bp_1, systolic_bp_2, systolic_bp_3, systolic_bp_4),
        na.rm = TRUE
      )
    )
  )

```

```{r final data sample}
final_df <- nhanes_df %>%
  filter(
    !is.na(age),
    !is.na(gender),
    !is.na(avg_sbp)
  )

```

```{r hypertension categories}
final_df <- final_df %>%
  mutate(
    sbp_category = case_when(
      avg_sbp < 120 ~ "Normal",
      avg_sbp >= 120 & avg_sbp < 130 ~ "Elevated",
      avg_sbp >= 130 & avg_sbp < 140 ~ "Stage 1 hypertension",
      avg_sbp >= 140 ~ "Stage 2 hypertension"
    )
  )

#ordering hypertension categories
final_df <- final_df %>%
  mutate(
    sbp_category = factor(
      sbp_category,
      levels = c(
        "Normal",
        "Elevated",
        "Stage 1 hypertension",
        "Stage 2 hypertension"
      )
    )
  )
```
## Relationship between age, gender, and systolic blood pressure
The analysis began with four systolic blood pressure (SBP) measurements for each participant, which were combined to calculate a single average SBP. Participants with missing age, gender, or average SBP were then excluded, resulting in a final analytical sample of `r nrow(final_df)` participants. This represents `r round(100 * nrow(final_df) / nrow(nhanes_df), 1)`% of the original `r nrow(nhanes_df)` participants, with `r nrow(nhanes_df) - nrow(final_df)` participants excluded because they did not have complete information required for the analysis.

The age distributions before and after filtering were broadly similar. In both samples, younger participants accounted for a substantial share of the observations, followed by a decline in participant counts through early adulthood and a more consistent distribution across the older age groups. The final sample ranged from `r min(final_df$age, na.rm = TRUE)` to `r max(final_df$age, na.rm = TRUE)` years, with `r names(which.max(table(final_df$age_group)))` representing the largest age group at `r max(table(final_df$age_group))` participants.

The median age was `r median(nhanes_df$age, na.rm = TRUE)` years in the original sample and `r median(final_df$age, na.rm = TRUE)` years in the final sample. Similarly, participants younger than 20 accounted for `r round(mean(nhanes_df$age < 20, na.rm = TRUE) * 100, 1)`% of the original sample and `r round(mean(final_df$age < 20, na.rm = TRUE) * 100, 1)`% of the final sample. These comparisons indicate that although complete-case filtering reduced the number of participants, it did not substantially alter the overall age composition of the dataset. The similar shapes of the two age distributions provide visual support for this conclusion.

The average systolic blood pressure was classified into four categories: Normal, Elevated, Stage 1 hypertension, and Stage 2 hypertension. Across the final sample, the most common category was `r names(which.max(table(final_df$sbp_category)))`, with `r max(table(final_df$sbp_category))` participants. 
This provided a starting point for examining whether the distribution of blood pressure categories differed across age groups and between genders.
```{r hypertension plot 1}
ggplot(final_df, aes(x = sbp_category)) +
  geom_bar() +
  labs(
    title = "Distribution of Blood Pressure Categories",
    x = "SBP Category",
    y = "Number of Participants"
  ) +
  theme_minimal()

```

```{r Gender distribution}
ggplot(final_df, aes(x = gender)) +
  geom_bar() +
  labs(
    title = "Gender Distribution in the Final Sample",
    x = "Gender",
    y = "Number of Participants"
  ) +
  theme_minimal()

```
```{r age distribution plot}
ggplot(final_df, aes(x = age)) +
  geom_histogram(binwidth = 5) +
  labs(
    title = "Age Distribution in the Final Sample",
    x = "Age",
    y = "Number of Participants"
  ) +
  theme_minimal()

```

```{r hypertension categories across age and gender}
final_df <- final_df %>%
  mutate(
    age_group = case_when(
      age < 20 ~ "0–19",
      age >= 20 & age < 40 ~ "20–39",
      age >= 40 & age < 60 ~ "40–59",
      age >= 60 & age < 80 ~ "60–79",
      age >= 80 ~ "80+",
      TRUE ~ NA_character_
    ),
    age_group = factor(
      age_group,
      levels = c(
        "0–19",
        "20–39",
        "40–59",
        "60–79",
        "80+"
      )
    )
  )

ggplot(
  final_df,
  aes(x = age_group, fill = sbp_category)
) +
  geom_bar() +
  facet_wrap(~ gender) +
  labs(
    title = "Blood Pressure Categories Across Age and Gender",
    subtitle = "Distribution of average systolic blood pressure categories by age group",
    x = "Age Group",
    y = "Number of Participants",
    fill = "SBP Category"
  ) +
  scale_fill_manual(
    values = c(
      "Normal" = "#4CAF50",
      "Elevated" = "#FFC107",
      "Stage 1 hypertension" = "#FF9800",
      "Stage 2 hypertension" = "#D32F2F"
    )
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```
```{r summary of distribution}
# Create a summary for the storytelling text
sbp_story <- final_df %>%
  group_by(age_group, gender) %>%
  summarise(
    n = n(),
    normal_pct = mean(sbp_category == "Normal") * 100,
    elevated_pct = mean(sbp_category == "Elevated") * 100,
    stage1_pct = mean(sbp_category == "Stage 1 hypertension") * 100,
    stage2_pct = mean(sbp_category == "Stage 2 hypertension") * 100,
    higher_bp_pct = mean(
      sbp_category %in% c("Stage 1 hypertension", "Stage 2 hypertension")
    ) * 100,
    .groups = "drop"
  )
sbp_story %>%
  mutate(
    normal_pct = round(normal_pct, 1),
    elevated_pct = round(elevated_pct, 1),
    stage1_pct = round(stage1_pct, 1),
    stage2_pct = round(stage2_pct, 1),
    higher_bp_pct = round(higher_bp_pct, 1)
  ) %>%
  knitr::kable(
    col.names = c(
      "Age Group",
      "Gender",
      "N",
      "Normal (%)",
      "Elevated (%)",
      "Stage 1 (%)",
      "Stage 2 (%)",
      "Stage 1 + 2 (%)"
    ),
    caption = "Distribution of systolic blood pressure categories by age group and gender"
  )
highest_hypertension <- sbp_story %>%
  slice_max(higher_bp_pct, n = 1)

lowest_hypertension <- sbp_story %>%
  slice_min(higher_bp_pct, n = 1)
```

The visualization for blood pressure categories across age and gender showed how systolic blood pressure categories were distributed across the five age groups for each gender. The distribution was examined by comparing the relative size of the blood pressure categories within each age group. In particular, the proportion of participants in the higher blood pressure categories can be compared across the younger and older age groups.

The largest age group in the sample is `r names(which.max(table(final_df$age_group)))`, while the smallest is `r names(which.min(table(final_df$age_group)))`. These differences in group size are important when interpreting the visualization because the bars represent participant counts rather than percentages.

The table and visualization show how average systolic blood pressure categories are distributed across age groups and gender. Among the youngest participants (`r levels(final_df$age_group)[1]`), the Normal category represents a substantial share of participants. As age increases, the distribution shifts toward the higher blood pressure categories.

The highest combined percentage of Stage 1 and Stage 2 hypertension occurs among `r highest_hypertension$gender` participants in the `r highest_hypertension$age_group` age group, where `r round(highest_hypertension$higher_bp_pct, 1)`% are classified in one of these two categories. In contrast, the lowest combined percentage occurs among `r lowest_hypertension$gender` participants in the `r lowest_hypertension$age_group` age group, at `r round(lowest_hypertension$higher_bp_pct, 1)`%.

The visualization makes this pattern particularly visible: the green Normal sections are more prominent among younger participants, while the orange and red sections representing Stage 1 and Stage 2 hypertension become more prominent in older age groups. The table provides the corresponding percentages, allowing the distributions to be compared more directly across groups.

Overall, the results show a clear shift in the distribution of systolic blood pressure categories across age groups, with higher blood pressure categories becoming more prominent among older participants. 

The differences between females and males was also be examined within each age group. Among participants aged 0–19, females had a higher proportion in the Normal category than males (`r round(sbp_story$normal_pct[sbp_story$age_group == "0–19" & sbp_story$gender == "Female"], 1)`% versus `r round(sbp_story$normal_pct[sbp_story$age_group == "0–19" & sbp_story$gender == "Male"], 1)`%), while the combined proportion with Stage 1 or Stage 2 hypertension is lower among females (`r round(sbp_story$higher_bp_pct[sbp_story$age_group == "0–19" & sbp_story$gender == "Female"], 1)`% versus `r round(sbp_story$higher_bp_pct[sbp_story$age_group == "0–19" & sbp_story$gender == "Male"], 1)`%).

This difference was more noticeable in the 20–39 age group, where `r round(sbp_story$normal_pct[sbp_story$age_group == "20–39" & sbp_story$gender == "Female"], 1)`% of females are in the Normal category compared with `r round(sbp_story$normal_pct[sbp_story$age_group == "20–39" & sbp_story$gender == "Male"], 1)`% of males. The corresponding percentages for Stage 1 and Stage 2 hypertension combined are `r round(sbp_story$higher_bp_pct[sbp_story$age_group == "20–39" & sbp_story$gender == "Female"], 1)`% for females and `r round(sbp_story$higher_bp_pct[sbp_story$age_group == "20–39" & sbp_story$gender == "Male"], 1)`% for males.

The gender difference was smaller in the 40–59 and 60–79 age groups. For example, among participants aged 60–79, the proportion with Stage 1 or Stage 2 hypertension is `r round(sbp_story$higher_bp_pct[sbp_story$age_group == "60–79" & sbp_story$gender == "Female"], 1)`% among females and `r round(sbp_story$higher_bp_pct[sbp_story$age_group == "60–79" & sbp_story$gender == "Male"], 1)`% among males. However, the pattern changes again among participants aged 80 and older. In this group, females have a higher proportion classified with Stage 1 or Stage 2 hypertension (`r round(sbp_story$higher_bp_pct[sbp_story$age_group == "80+" & sbp_story$gender == "Female"], 1)`%) than males (`r round(sbp_story$higher_bp_pct[sbp_story$age_group == "80+" & sbp_story$gender == "Male"], 1)`%). In particular, Stage 2 hypertension accounts for `r round(sbp_story$stage2_pct[sbp_story$age_group == "80+" & sbp_story$gender == "Female"], 1)`% of females compared with `r round(sbp_story$stage2_pct[sbp_story$age_group == "80+" & sbp_story$gender == "Male"], 1)`% of males.

Taken together, the table shows that the gender differences are not consistent across all age groups. Males have a greater share of higher blood pressure categories in the younger age groups, while the difference narrows among participants aged 40–79 and reverses among those aged 80 and older, where females have a greater share of Stage 1 and Stage 2 hypertension.
